---
title: "MGT 6203 Project R Markdown"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the necessary libraries 

```{r echo=FALSE, warning=FALSE}
packages = c("dplyr", "ggplot2",
             "tidyverse", "mltools", 
             "data.table", "caret", "party", 
             "rpart", "rpart.plot", "e1071", 
             "caTools", "class", "randomForest", 
             "pROC", "plotROC")

## Now load or install & load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```


## Read in the necessary data files

```{r}
objects = read.csv('datasets/objects.csv')
degrees = read.csv('datasets/degrees.csv')
relationships = read.csv('datasets/relationships.csv')
acquisitions = read.csv('datasets/acquisitions.csv')
ipos = read.csv('datasets/ipos.csv')
```

## Data Pre-processing

We perform the following pre processing steps before performing EDA. 

1. The data set contains data that goes all the way back to the 1930s. For our 
analysis, we consider only records post the year 1980. 


```{r}
objects_filtered = objects %>%
  filter(as.Date(founded_at) >= "1980-01-01")
```

2. Filter records with status - 'Acquired', 'IPO' or 'Closed'. Success is defined
by 'Acquired' or 'IPO' while Failure is defined as 'Closed'. 


```{r}
objects_filtered =  objects_filtered %>% 
  filter(status %in% c("ipo", "acquired", "closed"))

objects_filtered$status[which(objects_filtered$status == "ipo")] = "acquired"

```


3. Filter records where the entity type is a 'Company'.

```{r}
objects_filtered = objects_filtered %>% 
  filter(entity_type == 'Company')
```


4. Some of the categories (example of categories: e-commerce, software, biotech,
etc.) have missing values. The data set already contains a group category called
'other'. 

Due to this, we eliminate missing values as the fact that placing them in the 
other category may lead to creating/imputing incorrect data. Example: A missing
value for an originally Software company might be placed into the other. 

```{r}
objects_filtered = objects_filtered %>% 
  filter(category_code != "")
```


5. We have missing values for the Countries as well. We will remove these data 
records so as to not provide inaccurate information to the model. 

```{r}
objects_filtered = objects_filtered %>% 
  filter(country_code != "")
```


6. Instead of including the twitter username for each company, we will be modeling
it as a Binary Variable indicating Social media presence/absence. 

```{r}
objects_filtered$twitter_username[objects_filtered$twitter_username!=""] = 1
objects_filtered$twitter_username[objects_filtered$twitter_username==""] = 0
objects_filtered$twitter_username = as.numeric(objects_filtered$twitter_username)
objects_filtered$twitter_presence = objects_filtered$twitter_username
objects_filtered$twitter_username = NULL
```

7. We convert the category and country into factor type

```{r}
objects_filtered$category_code = as.factor(objects_filtered$category_code)
objects_filtered$country_code = as.factor(objects_filtered$country_code)
```

8. The response variable as indicated earlier is a binary variable. Success 
is indicated by 1, while Failure is indicated by 0. 

```{r}
objects_filtered$status = ifelse(objects_filtered$status == "acquired", 1,0)
```

9. We quickly verify if there are any duplicate company IDs. We find all the 
company IDs are unique

```{r}
anyDuplicated(objects_filtered$id)
unique_companies = objects_filtered$id
```

10. We read in the relationships and degrees data and only filter based on the 
companies that exist in the objects_filtered data frame. We then merge them 
based on the person ID which acts as the foreign key. 

```{r}
relationships_filtered = relationships %>% 
  filter(relationship_object_id %in% unique_companies) 

degrees_filtered = degrees %>% 
  filter(object_id %in% relationships$person_object_id)

rel_degree_merged = left_join(relationships_filtered, degrees_filtered,
                              by = c("person_object_id" = "object_id"))
```

11. We perform some pre-processing to find the following four features - 

a. No. of Founders
b. No. of MBAs
c. No. of MSs
d. No. of PHDs

Please note that these values are calculated for the Top Level Management 
positions. Therefore, they may be on the lower side with respect to quantity as
compared to the whole company. 

```{r}
no_of_founders = rel_degree_merged %>% 
  filter(grepl('founder', title, ignore.case = TRUE)) %>% 
  group_by(relationship_object_id) %>%
  summarize(No_Of_Founders = length(unique(person_object_id)))

no_of_mbas = rel_degree_merged %>% 
  filter(degree_type == 'MBA') %>% 
  group_by(relationship_object_id) %>% 
  summarise(No_Of_MBAs = length(unique(person_object_id)))

no_of_ms = rel_degree_merged %>% 
  filter(degree_type == 'MS') %>% 
  group_by(relationship_object_id) %>% 
  summarise(No_Of_MSs = length(unique(person_object_id)))

no_of_phds = rel_degree_merged %>% 
  filter(degree_type == 'PhD') %>% 
  group_by(relationship_object_id) %>% 
  summarise(No_Of_PHDs = length(unique(person_object_id)))
```

12. We perform a left join based on our original objects_filtered data frame
and the above data frames. The ID here will refer to the company ID placed as 
a foreign key in the rel_degree_merged data frame. 

```{r}
objects_filtered = left_join(objects_filtered, no_of_founders, 
                             by = c("id" = "relationship_object_id"))
objects_filtered = left_join(objects_filtered, no_of_mbas, 
                             by = c("id" = "relationship_object_id"))
objects_filtered = left_join(objects_filtered, no_of_ms, 
                             by = c("id" = "relationship_object_id"))
objects_filtered = left_join(objects_filtered, no_of_phds, 
                             by = c("id" = "relationship_object_id"))

objects_filtered$No_Of_Founders[is.na(objects_filtered$No_Of_Founders)] = 1
objects_filtered$No_Of_MBAs[is.na(objects_filtered$No_Of_MBAs)] = 0
objects_filtered$No_Of_MSs[is.na(objects_filtered$No_Of_MSs)] = 0
objects_filtered$No_Of_PHDs[is.na(objects_filtered$No_Of_PHDs)] = 0
```

13. We then model the countries as to whether they belong to a top 5 country 
or not, top 5 indicated by the no. of start-ups emerging from that country. 

```{r}
top_5_countries = objects_filtered %>%
  group_by(country_code) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(5) %>%
  dplyr::select(country_code)

top_5_countries$is_top_5_country = 1
objects_filtered = left_join(objects_filtered, top_5_countries,
                             by = c("country_code" = "country_code"))
objects_filtered$is_top_5_country[is.na(objects_filtered$is_top_5_country)] = 0
```

14.  We consider only records with funding rounds greater than 0 because when we
consider companies with 0 funding rounds that turn out to be successful, they 
are self sustainable and the investors don't get an opportunity to invest in 
such companies. Keeping companies with 0 funding rounds would thus degrade our
prediction models. 

```{r}
objects_filtered = objects_filtered %>% filter(funding_rounds > 0)
```

15. We filter only the necessary columns from the objects_filtered data frame. 

```{r}
objects_filtered_final = objects_filtered %>% dplyr::select(category_code, 
                                                     is_top_5_country,
                                                     milestones, relationships, 
                                                     funding_rounds, 
                                                     funding_total_usd,
                                                     twitter_presence, 
                                                     No_Of_Founders, No_Of_MBAs,
                                                     No_Of_MSs, No_Of_PHDs,
                                                     status)
```

## Exploratory Data Analysis

### Bar Chart of No. of Companies in each Category 

```{r}
no_of_companies_vs_category = objects_filtered %>% 
  filter(category_code != "") %>% 
  group_by(category_code) %>% 
  summarise(Total = n()) %>% 
  arrange(desc(Total))

p<-ggplot(data=head(no_of_companies_vs_category, 20), aes(x=reorder(category_code, -Total), y=Total, fill=category_code)) +
  geom_bar(stat="identity", position='dodge') + 
  labs(title = "Number of Companies in Each Category (Top 20 shown)", x = "", y = "Number of Companies") + scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=Total), position=position_dodge(width=0.9), vjust=-0.25) +
  guides(fill="none") + 
  theme_minimal()
p
```

### Bar Chart of No. of Companies Acquired in each Category

```{r}
no_of_companies_acquired_vs_category = objects_filtered %>% 
  filter(category_code != "") %>% 
  group_by(category_code, status) %>%
  summarise(acquired = sum(status == 1), closed = sum(status == 0)) %>% 
  group_by(category_code) %>% 
  filter(status == 1) %>% 
  dplyr::select(acquired) %>%
  arrange(desc(acquired))

p<-ggplot(data=head(no_of_companies_acquired_vs_category, 20), 
          aes(x=reorder(category_code, -acquired), y=acquired, 
              fill=category_code)) +
  geom_bar(stat="identity", position='dodge') + 
  labs(title = "Number of Companies Acquired in Each Category (Top 10 Shown)", 
       x = "Category", y = "Number of Companies Acquired") + 
  scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=acquired), position=position_dodge(width=0.9), 
            vjust=-0.25) + guides(fill="none") + theme_minimal()
p
```

### Acquisition Rate for each Category (Minimum of 40 acquisitions)

```{r}
acquisition_rate_vs_category = merge(no_of_companies_vs_category, 
                                     no_of_companies_acquired_vs_category, 
                                     by='category_code')

acquisition_rate_vs_category = acquisition_rate_vs_category %>% 
  mutate(acquired_rate = round((acquired/Total)*100)) %>% 
  filter(acquired >= 40) %>% 
  arrange(desc(acquired_rate))

p<-ggplot(data=head(acquisition_rate_vs_category, 20),
          aes(x=reorder(category_code, -acquired_rate),
              y=acquired_rate, fill=category_code)) +
  geom_bar(stat="identity", position='dodge') + 
  labs(title = "Success Rate per Category (Top 20 Shown)", x = "", 
       y = "Success Rate") + 
  scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=acquired_rate), position=position_dodge(width=0.9),
            vjust=-0.25) +
  guides(fill="none") +
  theme_minimal()
p
```
### Relationships vs Acquisition Rate

```{r}
relationships_acquired_status = objects_filtered %>% 
  group_by(gr=cut(relationships,breaks= c(0,2,4,6,8,10,20,30,40,50,100,Inf)
, right=FALSE),status) %>% 
     summarise(n= n()) %>%
     arrange(as.numeric(gr)) %>%
  filter(status == 1) %>%
  arrange(desc(n))


relationships_total_status = objects_filtered %>% 
  group_by(gr=cut(relationships,breaks= c(0,2,4,6,8,10,20,30,40,50,100, Inf)
, right=FALSE)) %>% 
     summarise(Total= n()) %>%
     arrange(as.numeric(gr))

relationships_acquired_rate = merge(relationships_total_status, 
                                    relationships_acquired_status, by='gr')

relationships_acquired_rate = relationships_acquired_rate %>% 
  mutate(acquired_rate = round((n/Total)*100))

p<-ggplot(data=head(relationships_acquired_rate, 10), aes(x=gr, y=acquired_rate,
                                                          fill=gr)) +
  geom_bar(stat="identity", position='dodge') + 
  labs(title = "Success Rate vs Number of Relationships", x = "Relationships", 
       y = "Success Rate") + 
  scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=acquired_rate), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  guides(fill="none") +
  theme_minimal()
p
```

### Acquisitions per Founding year 

```{r}
objects_filtered = objects_filtered %>% 
  mutate(Year = format(as.Date(objects_filtered$founded_at, "%Y-%m-%d"), "%Y"))

acquisitions_per_year = objects_filtered %>% 
  group_by(Year, status) %>% summarise(Acquired = sum(status == 1)) %>% 
  filter(status == 1) %>% arrange(Year)

companies_founded_per_year = objects_filtered %>% 
  group_by(Year) %>% 
  summarise(Total = n())

founded_acquisition_rate = merge(acquisitions_per_year, 
                                 companies_founded_per_year, by = 'Year') %>%
  mutate(acquired_rate = round((Acquired/Total)*100))

p <- ggplot(head(founded_acquisition_rate, 34), aes(x=Year, y=Total, group = 1)) +
   geom_point() + geom_line()+ scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=Total), position=position_dodge(width=0.9), vjust=-0.25) + 
  labs(title = "Companies Founded over the Years", x = "Founding Year", y = "Total No. of Companies")

q <- ggplot(head(founded_acquisition_rate, 34), aes(x=Year, y=acquired_rate, group = 1)) +
   geom_point() + geom_line()+ scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=acquired_rate), position=position_dodge(width=0.9), vjust=-0.25) + 
  labs(title = "Success Rate over the Years", x = "Founding Year", y = "Success Rate")
  
par(mfrow=c(2,1))
plot(p)
plot(q)
```


### Funding Rounds vs Success Rate

```{r}
funding_rounds_acquired_status = objects_filtered %>% 
  group_by(gr=cut(funding_rounds,breaks= c(1,2,3,4,5,6,7,8,9,10,11,12,13)
, right=FALSE),status) %>% 
     summarise(n= n()) %>%
     arrange(as.numeric(gr)) %>%
  filter(status == 1) %>%
  arrange(desc(n))


funding_rounds_total_status = objects_filtered %>% 
  group_by(gr=cut(funding_rounds,breaks=  c(1,2,3,4,5,6,7,8,9,10,11,12,13)
, right=FALSE)) %>% 
     summarise(Total= n()) %>%
     arrange(as.numeric(gr))


funding_rounds_acquired_rate = merge(funding_rounds_total_status, funding_rounds_acquired_status, by='gr')

funding_rounds_acquired_rate = funding_rounds_acquired_rate %>% 
  mutate(acquired_rate = round((n/Total)*100))

p<-ggplot(data=head(funding_rounds_acquired_rate, 12), aes(x=gr, y=acquired_rate, fill=gr)) +
  geom_bar(stat="identity", position='dodge') + 
  labs(title = "Success Rate vs Funding rounds", x = "Funding Rounds", y = "Success Rate") + 
  scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=acquired_rate), position=position_dodge(width=0.9), vjust=-0.25) +
  guides(fill="none") +
  theme_minimal()
p
```

### Companies Found Closed Year

```{r}
objects_filtered = objects_filtered %>% 
  mutate(Closed_Year = format(as.Date(objects_filtered$closed_at, "%Y-%m-%d"), 
                              "%Y"))

closed_year = objects_filtered %>% 
  filter(Closed_Year >= '1980') %>% 
  group_by(Closed_Year) %>% count()

nrow(closed_year)

ggplot(head(closed_year, 16), aes(x=Closed_Year, y=n, group = 1)) +
   geom_point() + geom_line()+ scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25) + 
  labs(title = "No. of Companies closed over the years", x = "Closing Year",
       y = "No. of Companies")
```

### Companies Found Acquisition Year

```{r}
acquisitions = acquisitions %>% mutate(Acquired_Year = format(as.Date(acquisitions$acquired_at, "%Y-%m-%d"), "%Y"))
ipos = ipos %>% mutate(IPO_Year = format(as.Date(ipos$public_at, "%Y-%m-%d"), "%Y"))

acquisitions_grouped = acquisitions %>% group_by(Acquired_Year) %>% count()
ipos_grouped = ipos %>% group_by(IPO_Year) %>% count()

acquisition_ipo_joined = left_join(acquisitions_grouped, ipos_grouped, by=c("Acquired_Year"="IPO_Year"))

acquisition_ipo_joined$Count = rowSums(acquisition_ipo_joined[,c("n.x", "n.y")], na.rm=TRUE)

ggplot(head(acquisition_ipo_joined, 33), aes(x=Acquired_Year, y=Count, group = 1)) +
   geom_point() + geom_line()+ scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  geom_text(aes(label=Count), position=position_dodge(width=0.9), vjust=-0.25) + 
  labs(title = "No. of Companies acquired/ipo over the years", x = "Acquired/IPO Year", y = "No. of Companies")
```

## Train, Validation and Test Split

We One hot encode the categorical variables present in the data set. 

We split our data set into Train (70%), Validation (15%) and Test (15%) sets. 
The train.csv, validation.csv and test.csv have already been written into the
datasets/ folder beforehand and therefore the 3 write.csv lines are commented 
out. 

Please head over to the file team043_r_markdown_models.Rmd to view the code
for building the ML Models. 

```{r}
newdata <- one_hot(as.data.table(objects_filtered_final))

set.seed(2)
train.index <- createDataPartition(newdata$status, p = .7, list = FALSE)
train <- newdata[ train.index,]
test  <- newdata[-train.index,]

set.seed(2)
validation.index <- createDataPartition(test$status, p = .5, list = FALSE)
validation = test[validation.index, ]
final_test = test[-validation.index, ]

# write.csv(train, 'datasets/train.csv')
# write.csv(validation, 'datasets/validation.csv')
# write.csv(final_test, 'datasets/test.csv')

```




